---
title: "ESM 244 Lab 4: Nonlinear Regression"
author: "Hanna Buechi"
date: "1/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Lab Week 4:

- Nonlinear least squares
- Panel regression level 0 example
- Shiny example with navbarPage

###Part 0. Load packages
```{r}
library(tidyverse)
library(Ecdat)
library(plm)
library(lmtest)
library(car) #for plotting
```

###Part 1. Nonlinear least squares (logistic growth fitting)

a. Load dataset CellGrowth.csv. Create a scatterplot.

```{r data_coefficients}

CellGrowth <- read_csv("CellGrowth.csv")

# Look at it

ggplot(CellGrowth, aes(x = Time, y = CellCount)) +
  geom_point()

# non-linear: Nt = A / (1 + Be^-(rt)), B = (K-N0)/N0 --> R will help us find A, B, and r parameters via Gauss-Newton convergence

# Predictor = time, outcome = cell count
# Find estimates for logistic growth
# Looking at ggplot, K = 3,700
# N0 = 2000

BEstimate <- (3700-2000)/2000

# estimate r: isolate explonential phase of growth, transform to semi-log plot (ln(y) and t plotted), slope of line is r --> 0 - 4 hrs based on ggplot

# Now we are going to estimate r (the growth rate)

lag_phase <- CellGrowth %>% 
  filter(Time < 4)

# ln(y) = rt

ggplot(lag_phase, aes(x = Time, y = log(CellCount))) +
  geom_point()

lm(log(CellCount) ~ Time, data = lag_phase)

# r = 0.1121

```

Find model A, B, and r using Gauss-Newton method for iterative convergence (non linear least squares):

Use the nls() function:

```{r}

cell_fit <- nls(CellCount ~ A/(1 + B*exp(-r*Time)),
                start = list(A = 3700, B = 0.85, r = 0.1121),
                data = CellGrowth,
                trace = TRUE) # trace back how R got these values

# values on the left are sq. residuals associated with the model until convergence

cell_fit

# Achieved convergence tolerance: 6.512e-06 --> residuals would only change by this factor if R tried again

# Store coefficients to be able to plot
A <- coef(cell_fit)[1] # first coefficient reported in cell_fit model is A
B <- coef(cell_fit)[2]
r <- coef(cell_fit)[3]

# Create a new sequence of values
time_seq <- seq(0,20,length = 100) # sequence of 1 to 20 with 100 increments

# Plug that new sequence into my model with the paragmeters A, B, and r that we found:

cell_pred <- A/(1+B*exp(-r*time_seq))

# Bind that together with the time series data for plotting:
pred_df <- data.frame(time_seq, cell_pred)

# Create a graph with original data and our model predictions:
ggplot(CellGrowth, aes(x = Time, y = CellCount)) +
  geom_point(color = "blue", size = 3) +
  geom_line(data = pred_df, aes(x = time_seq, y = cell_pred), color = "orange", size = 1)

# Watch Miles McVain's talk about magic in R

```

Panel Regression - First Shot

What is the effect of cigarette price on cigarette purchases?

State: our entity variables
Year: our time variable
packpc (pack per cap): our dependent variable
avgprs (average price): our predictor variable of interest

```{r wranlge}

cigs_panel <- Cigarette %>% 
  select(state, year, packpc, avgprs)

ggplot(cigs_panel, aes(x = avgprs, y = packpc, group = state)) +
  geom_point(aes(color = state)) +
  geom_smooth(method = "lm", aes(color = state), se = FALSE, size = 0.2) # se = FALSE excludes SE of lm

# overall across states, avg price affects pack purchases in similar ways

# What's happening over time?

ggplot(cigs_panel, aes(x = year, y = packpc, group = state)) +
  geom_point(aes(color = state)) +
  geom_smooth(method = "lm", aes(color = state), se = FALSE, size = 0.2)

# Trend through time is consistent, too

```

Entity fixed effects model:

```{r}

cigs_model <- plm(packpc ~ avgprs,
                  data = cigs_panel,
                  index = c("state","year"),
                  model = "within") # entity fixed effects model

cigs_model # the only coefficient shows up is the price one - this is interpreted the same way as for linear model coefficients: "if all else is held equal, or taking entity fixed effects into account, the number of packs bought decreases by 0.33 for every dollar increase in price"


# robust standard error

coeftest(cigs_model, vcov. = vcovHC(cigs_model, type = "HC1")) # variables are correlated with each other, error associated

# coefficient has changed a little bit and standard error is larger
# SE: 0.012974
```


Add time fixed effects:

```{r}

# same code as cigs_model

et_cigs_model <- plm(packpc ~ avgprs,
                  data = cigs_panel,
                  index = c("state","year"),
                  model = "within", # still have entity fixed effects
                  effect = "twoways") # time fixed effects

et_cigs_model

# coefficient is a little larger
# but has this improved the model
# pFtest compares models with null hypothesis that model is not improved adding time fixed effects

pFtest(et_cigs_model, cigs_model) # "more complete" model comes first
# reject the null hypothesis! p = 7.835e-10

```















